---
title: "Douglas Neumann Visualizations"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: spacelab
    highlight: kate
    
---

```{r setup, include=FALSE}
library(flexdashboard)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning  = FALSE)

```

 <font size="+6"><strong><center><em> Biden & Trump Twitter </em></center></strong></font>
 
<font size="+2">A key battleground for the 2020 Preseidential election was fought on Twitter. While the standard routine of giving stump speeches, shaking hands, and kissing babies was severely restricted due to the outbreak of the COVID-19 Pandemic, messaging on social media became even more important.Twitter became a key battleground for that messaging. The data came from tweets containing the use of #JoeBiden and #DonaldTrump. The following visualizations show the story of the 2020 election as it was fought in the trenches of twitter.</font>

```{r}
biden <- read.csv("C://Users//12607//Desktop//final_Project//biden_short.csv")
trump <- read.csv("C://Users//12607//Desktop//final_Project//trump_short.csv")
```

```{r}
library(ggplot2)
library(dplyr)
library(lubridate)
theme_set(theme_minimal())
```

```{r}
biden_day_count <- biden %>%
          group_by(created_at = as.Date(created_at)) %>%
          summarise(comments = n())
trump_day_count <- trump %>%
          group_by(created_at = as.Date(created_at)) %>%
          summarise(comments = n())
colnames(biden_day_count) <- c("date", "biden_count")
colnames(trump_day_count) <- c("date", "trump_count")

```

```{r}
combined_count <- biden_day_count %>% right_join(trump_day_count, by=c("date"))
fin_comb_count <- combined_count[-c(4), ] 
fin_biden_day_count <- biden_day_count[-c(4), ] 
fin_trump_day_count <- trump_day_count[-c(4), ]

```

```{r}
bll<- biden[-which(biden$lat == ""), ]
tll<- trump[-which(trump$lat == ""), ]
# bwl.1<- biden[-which(biden$country == "United States of America"), ]
# bwl.2<- bwl.1[-which(bwl.1$country == "United States"), ]
# bwl<- bwl.2[-which(bwl.2$country == ""), ]
# 
# twl.1<- trump[-which(trump$country == "United States of America"), ]
# twl.2<- twl.1[-which(twl.1$country == "United States"), ]
# twl<- twl.2[-which(twl.2$country == ""), ]

```

```{r}
biden_19_20 = biden[c(73384:75765),]
trump_19_20 = trump[c(124383:128243),]
```

```{r}
biden_19_20$likes <- as.numeric(as.character(biden_19_20$likes))
biden_19_20$user_followers_count <- as.numeric(as.character(biden_19_20$user_followers_count))

trump_19_20$likes <- as.numeric(as.character(trump_19_20$likes))
trump_19_20$user_followers_count <- as.numeric(as.character(trump_19_20$user_followers_count))
```


```{r}
library("cowplot")
library("googleway")
library("ggplot2")
library("ggrepel")
library("ggspatial")

library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
```

```{r}
bll$long <- as.numeric(as.character(bll$long))
bll$lat <- as.numeric(as.character(bll$lat))

tll$long <- as.numeric(as.character(tll$long))
tll$lat <- as.numeric(as.character(tll$lat))
```

```{r}
### Begin network vis code
library(rtweet)
library(igraph)
library(hrbrthemes)
library(ggraph)
library(tidyverse)
```

```{r}
biden_19_20$likes <- as.numeric(as.character(biden_19_20$likes))
biden_19_20$user_followers_count <- as.numeric(as.character(biden_19_20$user_followers_count))

trump_19_20$likes <- as.numeric(as.character(trump_19_20$likes))
trump_19_20$user_followers_count <- as.numeric(as.character(trump_19_20$user_followers_count))
```

```{r}
filter(biden_19_20, retweet_count > 0) %>% 
  select(user_name, city) %>%
  unnest(city) %>% 
  filter(!is.na(city)) %>% 
  graph_from_data_frame() -> rt_g

filter(trump_19_20, retweet_count > 0) %>% 
  select(user_name, city) %>%
  unnest(city) %>% 
  filter(!is.na(city)) %>% 
  graph_from_data_frame() -> rt_gt
```

```{r}
filter(biden_19_20, retweet_count > 10) %>% 
  select(user_name, city) %>%
  unnest(city) %>% 
  filter(!is.na(city)) %>% 
  graph_from_data_frame() -> rt_g2

filter(trump_19_20, retweet_count > 10) %>% 
  select(user_name, city) %>%
  unnest(city) %>% 
  filter(!is.na(city)) %>% 
  graph_from_data_frame() -> rt_gt2
```

```{r}
V(rt_g2)$node_label <- unname(ifelse(degree(rt_g2)[V(rt_g2)] > 2, names(V(rt_g2)), "")) 
V(rt_g2)$node_size <- unname(ifelse(degree(rt_g2)[V(rt_g2)] > 2, degree(rt_g2), 0)) 

V(rt_gt2)$node_label <- unname(ifelse(degree(rt_gt2)[V(rt_gt2)] > 2, names(V(rt_gt2)), "")) 
V(rt_gt2)$node_size <- unname(ifelse(degree(rt_gt2)[V(rt_gt2)] > 2, degree(rt_gt2), 0)) 

### End Network vis code
```



```{r}
### Begin Text vis code

library(readr)
library(dplyr)
library(e1071)
library(mlbench)
 
#Text mining packages
library(tm)
library(SnowballC)
library("wordcloud")
library("RColorBrewer")
```

```{r}
biden_19_20$tweet <- as.character(biden_19_20$tweet)
#glimpse(biden_19_20)
# Create corpus
corpus = Corpus(VectorSource(biden_19_20$tweet))
# Look at corpus
#corpus[[1]][1]

trump_19_20$tweet <- as.character(trump_19_20$tweet)

# Create corpus
corpus_t = Corpus(VectorSource(trump_19_20$tweet))

```

```{r}
#Conversion to Lowercase
corpus = tm_map(corpus, PlainTextDocument)
corpus = tm_map(corpus, tolower)
 
#Removing Punctuation
corpus = tm_map(corpus, removePunctuation)

#Remove stopwords
corpus = tm_map(corpus, removeWords, c("que", stopwords("english")))
corpus = tm_map(corpus, removeWords, c("this", stopwords("english")))
corpus = tm_map(corpus, removeWords, c("and", stopwords("english")))
corpus = tm_map(corpus, removeWords, c("will", stopwords("english")))
#corpus = tm_map(corpus, removeWords, c("are", stopwords("english")))

# Stemming
corpus = tm_map(corpus, stemDocument)
 
# Eliminate white spaces
corpus = tm_map(corpus, stripWhitespace)

#trump code begins
#Conversion to Lowercase
corpus_t = tm_map(corpus_t, PlainTextDocument)
corpus_t = tm_map(corpus_t, tolower)
 
#Removing Punctuation
corpus_t = tm_map(corpus_t, removePunctuation)

#Remove stopwords
corpus_t = tm_map(corpus_t, removeWords, c("que", stopwords("english")))
corpus_t = tm_map(corpus_t, removeWords, c("this", stopwords("english")))
corpus_t = tm_map(corpus_t, removeWords, c("and", stopwords("english")))
corpus_t = tm_map(corpus_t, removeWords, c("will", stopwords("english")))
#corpus = tm_map(corpus, removeWords, c("are", stopwords("english")))

# Stemming
corpus_t = tm_map(corpus_t, stemDocument)
 
# Eliminate white spaces
corpus_t = tm_map(corpus_t, stripWhitespace)
#corpus[[1]][1] 
```

```{r}
DTM <- TermDocumentMatrix(corpus)
mat <- as.matrix(DTM)
f <- sort(rowSums(mat),decreasing=TRUE)
dat <- data.frame(word = names(f),freq=f)

DTMt <- TermDocumentMatrix(corpus_t)
matt <- as.matrix(DTMt)
ft <- sort(rowSums(matt),decreasing=TRUE)
dat_t <- data.frame(word = names(ft),freq=ft)
```

```{r}
bid_words <- head(dat, 20)
trump_words <- head(dat_t, 20)
### end Text vis code
```


```{r}
rm(trump)
rm(biden)
```

Column {.tabset .tabset-fade}
-----------------------------------------------------------------------

### The Data

The data for this analysis comes from twitter and via kaggle. It was initally was over a 25 day period starting in the middle of October and going a few days past the elections. Due to the sheer magnitude of the data it was redcuced in scope to make processing possible.

```{r}
bll_head <- head(bll, 5)
bll_head
```



### Time Series - Comparison

```{r}
# Area plot
ggplot(fin_comb_count) + 
  geom_area(aes(x=date, y= trump_count), 
             color = "darkred", fill = "darkred") + 
  geom_area(aes(x=date, y= biden_count), 
             color = "darkblue", fill = "darkblue")+labs(title="Use of #JoeBiden (Blue) and #DonaldTrump (Red)", 
         x="Date", y = "Number of Tweets Containing Each #", subtitle = "Final 3 days Leading to the Election", caption = "")
        

```

This is the time series explanation

### Time Series - Biden
```{r}
# Basic line plot Biden
ggplot(data = fin_biden_day_count, aes(x = date, y = biden_count))+
  geom_line(color = "darkblue", size = 2)+
  labs(title="Use of #JoeBiden", subtitle = "Final 3 days Leading to the Election",
         x="Date", y = "Number of Tweets Containing the #")
```  


### Time Series - Trump
```{r}
# Basic line plot Trump
ggplot(data = fin_trump_day_count, aes(x = date, y = trump_count))+
  geom_line(color = "red", size = 2)+labs(title="Use of #DonaldTrump", subtitle = "Final 3 days Leading to the Election",
         x="Date", y = "Number of Tweets Containing the #")
```  


### Exploratory - Biden 
```{r}
#pairs plot
biden_19_20%>%
  select(!c(X)) %>% 
  select_if(function(x){is.numeric(x)})%>%
  filter((retweet_count < 500 | user_followers_count < 5000)) %>% 
  pairs(main="Biden Comparison Plot")
```


### Exploratory - Trump 
```{r}
#pairs plot
biden_19_20%>%
  select(!c(X)) %>% 
  select_if(function(x){is.numeric(x)})%>%
  filter((retweet_count < 500 | user_followers_count < 5000)) %>% 
  pairs(main="Trump Comparison Plot")
```



### Geographical - Biden
```{r}

world <- ne_countries(scale = "medium", returnclass = "sf")


ggplot(data = world) +
    geom_sf() +
    geom_point(data = bll, aes(x = long, y = lat), size = .1, 
        shape = 23, fill = "darkblue")+labs(title="World Map of #JoeBiden ", subtitle = "Tweeted from 7:00 pm to 8:00 pm on Election Night", x="", y = "")


```



### Geographical - Trump
```{r}

world <- ne_countries(scale = "medium", returnclass = "sf")


ggplot(data = world) +
    geom_sf() +
    geom_point(data = tll, aes(x = long, y = lat), size = .1, 
        shape = 23, fill = "darkred")+labs(title="World Map of #DonaldTrump ", subtitle = "Tweeted from 7:00 pm t0 8:00 pm on Election Night", x="", y = "")


```

### Network - Biden
```{r}
ggraph(rt_g2, layout = 'linear', circular = TRUE) + 
  geom_edge_arc(edge_width=0.125, aes(alpha=..index..,start_cap = label_rect(node1.name),end_cap = label_rect(node2.name))) +
  # geom_node_label(aes(label=node_label, size=node_size),
  #                 label.size=10, fill="#ffffff66", segment.colour="springgreen",
  #                 color="slateblue", repel=TRUE, family=font_rc, fontface="bold") +
  geom_node_text(aes(label=node_label))+
  coord_fixed() +
  scale_size_area(trans="sqrt") +
  labs(title="Biden Twitter City Relationships", subtitle="Nodes = (user_name and city). Darker Edges == More Connections") +
  theme_graph(base_family=font_rc) 
```

### Network - Trump
```{r}
ggraph(rt_gt2, layout = 'linear', circular = TRUE) + 
  geom_edge_arc(edge_width=0.125, aes(alpha=..index..,start_cap = label_rect(node1.name),end_cap = label_rect(node2.name))) +
  # geom_node_label(aes(label=node_label, size=node_size),
  #                 label.size=10, fill="#ffffff66", segment.colour="springgreen",
  #                 color="slateblue", repel=TRUE, family=font_rc, fontface="bold") +
  geom_node_text(aes(label=node_label))+
  coord_fixed() +
  scale_size_area(trans="sqrt") +
  labs(title="Trump Twitter City Relationships", subtitle="Nodes = (user_name and city). Darker Edges == More Connections") +
  theme_graph(base_family=font_rc) 
```

### Text - Wordcloud Biden
```{r}
set.seed(10)
wordcloud(words = bid_words$word, freq = bid_words$freq, random.order=FALSE, rot.per=0.30, colors=brewer.pal(8, "Dark2"))
```

### Text - Wordcloud Trump
```{r}
set.seed(10)
wordcloud(words = trump_words$word, freq = bid_words$freq, random.order=FALSE, rot.per=0.30, colors=brewer.pal(8, "Dark2"))
```


### Text - Bar Grapah Biden
```{r}
# Minimal theme + blue fill color
p<-ggplot(data=bid_words, aes(x=word, y=freq)) +
  geom_bar(stat="identity", fill="darkblue")+
  theme_minimal()
p+ coord_flip()+ labs(title="Most Used Words in Biden Twitter", subtitle = "Tweeted from 7:00 pm to 8:00 pm on Election Night",
         x="Word", y = "Count")
```


### Text - Bar Graph Trump
```{r}

p<-ggplot(data=trump_words, aes(x=word, y=freq)) +
  geom_bar(stat="identity", fill="darkred")+
  theme_minimal()
p+ coord_flip()+labs(title="Most Used Words in Trump Twitter",subtitle = "Tweeted from 7:00 pm to 8:00 pm on Election Night", 
         x="Word", y = "Count")

```
